---
title: "p8105_hw2_jx2371"
author: "Jingyu Xu"
date: "2018/10/5"
output: github_document
---

Firstly，I set a globle option which hides the message and warning information.
```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

#Problem 1
##read and clean the dataset
Firstly, a code chunck is writted to import the dataset
```{r import data1}
library(tidyverse)
NYC_Transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

Then, a code chunck is written to clean the data according to the reqirement. I name the manipulated dataset NYC_Transit_data1
```{r data cleaning}
NYC_Transit_data1=janitor::clean_names(NYC_Transit_data) %>%
  select(line:route11, entry, vending, entrance_type, ada)%>% #retain rows
  mutate(entry= ifelse(entry=="YES",TRUE,FALSE))#convert entry variable to logical variable
```

##describe the dataset 
The original dataset contains variables including "`r names(NYC_Transit_data)`"   
And the dataset we made some cleaning in the previous chunk contains variables including "`r names(NYC_Transit_data1)`"  

The original dataset contains variables including "`r names(NYC_Transit_data)`"   

The data cleaning steps is described as following:
1)use janitor::clean_names to transfer the variable names into lower snake case.    
2)retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.    
3)convert the entry variable from character(YES VS NO) to a logical variable using ifelse function. 

The dimension of the resulting dataset is `r dim(NYC_Transit_data1)[1]`&times;`r dim(NYC_Transit_data1)[2]`  

I think the data is untidy, because some of the column names（route1-route11） are not names of variables, but still values of a variable. As a result of that, each row contains more than a single observation.

##answers to dataset related questions
1.There are `r nrow(distinct(NYC_Transit_data1, line, station_name))` distinct stations. 

2.There are `r length(which(NYC_Transit_data1$ada==TRUE))`staion that are ADA compliant.  

3.The proportion of station entrance/exits without vending allow entrance is `r   length(which(NYC_Transit_data1$vending=="NO"))/length(NYC_Transit_data1$vending)` 

4.a code chunk is wrote to reformat the data so that the route number and route name are distinct
```{r}
gather(NYC_Transit_data1, key = route_number, value = route_name, route1:route11)
```

5.Among the distinct stations, how many station serve the A train?


6. Of the stations that serves A train, how many are ADA compliant?


#Problem 2
## Mr. Trash Wheel Sheet
Firstly, import the excel file and specify the Mr. Trash Wheel Sheet. I omit the note column by range function. 
```{r}
library(readxl)
trash_wheel= read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) 
```

Then，a code chunk is written to use reasonable names of the Mr. Trash Sheet. I delete the unit of the variable name to make it more readable.
```{r}
trash_wheel= janitor::clean_names(trash_wheel)#clean up the variables to the lower snake style
names(trash_wheel)#view the new variable name
trash_wheel=select(trash_wheel, volume=volume_cubic_yards, weight=weight_tons, everything())
names(trash_wheel)# view the modified name
```

Now, a code chunk is written to omit the rows that do not include dumpster-specific data
```{r}
trash_wheel_clean=filter(trash_wheel, !is.na(dumpster))
```

Then, we omit the rows that do not include dumpster-specific data and round the number of sports balls to the nearest integer and convert the result to an integer variable.
```{r}
trash_wheel_clean=filter(trash_wheel,!is.na(dumpster))%>%
  mutate(sports_balls=round(sports_balls,digits = 0))%>%
  mutate(sports_balls=as.integer(sports_balls))
```

## precipitation data for 2016 and 2017
Now, we read the precipitatin data for 2016 and 2017. Also we use janitor::clean_names to make the name more reasonable
```{r}
precipitation_2016= read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation", range= cell_rows(2:15))
precipitation_2017= read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range= cell_rows(2:15))
precipitation_2016= janitor::clean_names(precipitation_2016)
precipitation_2017= janitor::clean_names(precipitation_2017)
```

omit the rows without precipitation data and add a variable year
```{r precipitation_2016}
precipitation_2016_clean=filter(precipitation_2016,!is.na(month)) %>%
mutate(year=2016)
```

```{r}
precipitation_2017_clean=filter(precipitation_2017,!is.na(month)) %>%
mutate(year=2017)
```


now，combine the two datasets by month
```{r}
combine_data = left_join(precipitation_2016, precipitation_2017, by = "month")
mutate(combine_data, month=month.name)
```

## description of two datasets
And then caculate the median number of the sports balls in a dumpster in 2016
```{r}
A=filter(trash_wheel_clean, year==2016)
median(A$sports_balls)
```

The total preciitition in 2017 is `r sum(combine_data$total_2017)`


